### 2015-08-24 qdpm_paper.tex
Included some information of the QDPM hamiltonian and the Z3 local gauge symetry of the QDPM.

@@

    It became apparent in the 80's and early 90's that the Landau theory of symmetry breaking did not
    encompass all possible types of order possible in the groundstate of a condensed matter system. 
    In fact it was found that the symmetry of a system
    could increase as a system was cooled past a critical tempature and a phase transition occured
    that did not fit into Landau's theory. In these cases oder was found
    that could not be described with an order
    parameters \cite{wen_1990} which was a topological order does not fit in to Landau's
    theory of symmetry breaking. Though the discovery of topological order was originally thought
    to help understand superconductivity in the context of chiral spin systems \cite{HERE} it was
    not able to do so. However, due to the robust ground state degeneracy of systems exhibiting
    topological order they have become candidates for methods of storing and manipulating
    quantum information.  Also important to the field of quantum information, but interesting in
    its own right, is that topologically ordered systems exhibit fractal excitations.

@@

        \subsubsection{String operators}

@@

            The ground state of the $Z_3$ toric code on the torus is nine-fold degenerate because of
            the nine different topological sectors in this topology. One can move between the
            topological sectors with a nontrivial closed electric string which wraps around either
            the major axis or the minor axis of the torus. We will call the electric winding
            operator that acts around the major axis $W^e_y$ and around the minor axis $W^e_x$.
            A given topological sector can be distinguish 
            by acting around the minor and major axes with the magnetic string operator $W^m_y$ and
            $W^m_x$. All of the aforementioned winding operators commute with the Hamiltonian and
            therefore share the same eigenvalues. The operator $W^m_{x}$ ($W^m_{y}$) can 
            have three different values depending on how many electric strings enclose the minor
            (major) axis. Since there are three different
            possible values for each winding number around each axis there are nine distinct
            topological sectors.

@@

    \subsection{How to get local $Z_3$ symmetry in square lattice dimer model}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dimer Correlation Results}
    \subsection{Correlation functions}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimating the Gap}
    \subsection{Imaginary time vison correlations}

    \subsection{Imaginary time dimer correlations}

>>>>

\newcommand*{\hprs}{%
  \text{% change size in subscripts or superscripts
    \includegraphics[
      height=1.2ex,% adjust to suit
      valign=M,% center vertically
      raise=\fontdimen22\textfont2,% but raise it to the formula axis
    ]{Symbols/hrzprestar} }
}
\newcommand*{\hspr}{%
  \text{% change size in subscripts or superscripts
    \includegraphics[
      height=1.2ex,% adjust to suit
      valign=M,% center vertically
      raise=\fontdimen22\textfont2,% but raise it to the formula axis
    ]{Symbols/hrzstarpair} }
}

@@

    It became apparent in the 80's and early 90's that the Landau theory of symmetry breaking did not
    encompass all possible types of order that can be found in the groundstate of a condensed matter system. 
    In fact it was found that the symmetry of a system
    could increase as a system was cooled past a critical tempature and a phase transition occured
    that did not fit into Landau's theory. In these cases the odering 
    that could not be described with an order
    parameters \cite{wen_1990} was a topological order. Though the discovery of topological order was originally thought
    to help understand superconductivity in the context of chiral spin systems \cite{HERE} it was
    not able to do so. However, due to the robust ground state degeneracy of systems exhibiting
    topological order they have become candidates for methods of storing and manipulating
    quantum information.  Also important to the field of quantum information, but interesting in
    its own right, is that topologically ordered systems exhibit fractal excitations.

@@

@@

            The ground state of the $Z_3$ toric code on the torus is nine-fold degenerate because of
            the nine different topological sectors in this topology. One can move between the
            topological sectors with a nontrivial closed electric string wraps around either
            the major axis or the minor axis of the torus. We will call the electric winding
            operator that acts around the major axis $W^e_y$ and around the minor axis $W^e_x$.
            A given topological sector can be distinguish 
            by acting around the minor and major axes with the magnetic string operator $W^m_y$ and
            $W^m_x$. All of the aforementioned winding operators commute with the Hamiltonian and
            therefore share the same eigenvalues. The operator $W^m_{x}$ ($W^m_{y}$) can 
            have three different values depending on how many electric strings enclose the minor
            (major) axis. Since there are three different
            possible values for each winding number around each axis there are nine distinct
            topological sectors.

@@

    \subsection{Local $Z_3$ Gauge Symmetry}

        In the QDPM Ising degrees of freedom live on the links of the lattice.
        The operator $\hat{\sigma}^x_l$ acts on these links and has 
        eigenvalues $\sigma_l = \pm 1$, where $\sigma_l$ is the
        Ising variables of the $l^{\mathrm{th}}$ link on the lattice. The number operator for a give link is
        defined as $\hat{n}_l \equiv (1+\hat{\sigma}_x)/2$ and counts the number of dimers on link $l$.
        An exactl local gauge symetry exsists on the lattice due to the contraint that either one or
        four dimers touch any give vertex. In terms of $n_l$ we can define an operator that counts the
        number of dimers touching a given vertex $v$ as $n_v = \sum_{l\in v} n_l$.

        the local gauge transformation $G_v$ acting on a give vertex $v$ for the quantum dimer model
        can be expressed as
        %
        \begin{equation}
            \label{}
            G_v=e^{i \alpha (n_v - n_0)}
        \end{equation}
        %
        where $n_0$ is the constrained number of dimers touching a vertex, $n_0=1$ in the QDM model,
        The operator $G_v$ then acts in such a
        way that only the physical states, states satisfying the local constraint 
        are invariant. For the QDM $G_v =
        e^{i\alpha (0)}$ and therefore $G_v$ leaves a physical state unchanged for any
        $\alpha$ making the local gauge symetry $U(1)$. The local gauge transformation that leaves
        only configurations contaiing dimers and pentamers invarient is
        %
        \begin{equation}
            \label{eqn:gauge_trans}
            G_v = e^{i (\hat{n}_v -1) 2\pi/3}    
        \end{equation}
        %
        For the QDPM $\alpha$ is must be $2\pi/3$ due to the relaxed contraint of the QDM
        model that alows for pentamers. Relaxing the contraint of the QDM brings the local gauge
        symetry from $U(1)$ to $Z(3)$.

@@

        We work specificly at the RK point at wich the ground state wave function
        is an equal weighted superposition of fully packed dimer pentamer configurations
        %
        \begin{equation}
            \label{}
            \left| \Psi \right\rangle = \frac{1}{\sqrt{N}}\sum_{\{C\}} \left| C \right\rangle
            ,
        \end{equation}
        %
        where $\{C\}$ is the set of fully packed configurations. The Hilbert space can be defined
        using the gauge trasformation in Eq.~\ref{eqn:gauge_trans} as all $\left| C \right\rangle$
        satisfying $G \left| C \right\rangle = \left| C \right\rangle$.

        The QDPM Hamiltonian
        \begin{equation}
            H_{\mathrm{QDPM}} = H_{\mathrm{QDM}} + \mathrm{pentamer\ terms}
        \end{equation}

        The pentamer terms in the Hamiltonian can be constructed with the dynamics shown in
        Fig.~\ref{fig:pentamer_moves}.  
        \begin{figure}[htpb]
            \centering
            \includegraphics[width=0.8\linewidth]{pentamer_moves.pdf}
            \caption{a) Horizontal criation an anihilation, b) horizontal move, c) and d) are two
            differen possible diagonal moves in the same direction.}
            \label{fig:pentamer_moves}
        \end{figure}
        Each of the  moves shown in Fig.~\ref{fig:pentamer_moves} can be
        rotated to produce the symetry related counterpart (eg. vertical creation and anihilation). 
        As an example the creation and anhilation term in the Hamiltonian is
        \begin{multline}
        H_{\mathrm{pentamer\ creation}} = t_{\mathrm{create}}
            \sum_{\bstr}
            \left(
                \left|
                    \hprs
                \right\rangle
                \left\langle
                    \hspr
                \right|
                +
                \mathrm{h.c.}
            \right)
            \\
            - v_{\mathrm{create}}
            \sum_{\bstr}
                \left(
                    \left|
                        \hprs
                    \right\rangle
                    \left\langle
                        \hprs
                    \right|
                    +
                    \left|
                        \hspr
                    \right\rangle
                    \left\langle
                        \hspr
                    \right|
                \right)
        \end{multline}
        where the factors $t_{\mathrm{create}}$ and $v_{\mathrm{create}}$ are the coefficients of
        the penamer pair creation and anihilation respectivly. At the RK point
        $t/v=1$ for each all the petamer terms. 
        Writing the kinetic and potential energy terms as done above for all the dynamics shown in
        Fig.~\ref{fig:pentamer_moves} and all of their symetry related counterparts comprise the
        the pentamer terms of the Hamiltonian.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dimer Correlation Results}

@@

    \subsection{Correlation functions}

    \begin{equation}
        C_{\mathrm{dimer}} = \langle d_0 d_r \rangle - \langle d_0 \rangle   \langle d_r \rangle   
    \end{equation}

    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{spacial_dmr_cor.pdf}
        \caption{(Note: This figure was made using equal probabilities of plaquette filps and
        pentamer moves) Spacial correlation function of parralel dimers for the two different
        sublattices distiguesed by their color (red, green).}
        \label{fig:spacial_dmr_cor}
    \end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimating the Gap}
    \subsection{Imaginary time vison correlations}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{z2_origin_vison_time_cor.pdf}
        \caption{Imaginary time $z_2$ \textit{single} vison correlations at the origin in the QDM model (red) and QDPM
        model (green) on an $8\times$ lattice. For the QDPM the dimer/pentamer move fraction is $0.9$.}
        \label{fig:name}
    \end{figure}

    \subsection{Imaginary time dimer correlations}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{dimer_origin_time_cor.pdf}
        \caption{Imaginary time dimer correlations of a single dimer in the QDM model (red) and QDPM
        model (green) on an $8\times$ lattice. For the QDPM the dimer/pentamer move fraction is $0.9$. }
        \label{fig:}

    \end{figure}
